On-the-fly Algorithms for GKAT Equivalences \thanks{Identify applicable funding agency here. If none, delete this.}

### 2025-01-15 main.tex
proof, WIP

\newcommand\altxrightarrow[2][0pt]{\mathrel{\ensurestackMath{\stackengine%
  {\dimexpr#1-7.5pt}{\xrightarrow{\phantom{#2}}}{\scriptstyle\!#2\,}%
  {O}{c}{F}{F}{S}}}}
\newcommand{\transvia}[1]{
    \mathrel{\raisebox{-2px}{\(\altxrightarrow[-2px]{#1}\)}}
}
\newcommand{\transAcc}[2]{⇒_{#1} #2}

@@

\begin{theorem}[normalization perserves liveness]
    A state \(s ∈ S\) is dead (live) in \(S\) if and only if it is dead (live) in \(\norm(S)\).
\end{theorem}

@@

\begin{definition}[normalized bisimulation]\label{def:norm-bism}
    A relation \({≃} ⊆ S × U\) on two GKAT coalgebra \(S\) and \(U\) is a normalized bisimulation if and only if for any two states \(s ≃ u\), all the following holds:
    \begin{enumerate}
        \item\label{itm:acc-condition} for all \(α ∈ \At\), \(δ_{S}(s, α) = \accept ⟺ δ_{U}(u, α) = \accept\);
        \item\label{itm:rej-or-dead} if \(s\) reject \(α\) but \(u\) transitions to \(u'\), then \(u'\) is dead; similarly when \(u\) rejects \(α\);  
        \item\label{itm:transition-bisim} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', p)\), then \(s' ≃ u'\).
        \item\label{itm:transition-dead} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(t'\) are dead.
    \end{enumerate}
\end{definition}

\begin{lemma}[bisimulation between dead states]\label{thm:bisim-between-dead}
    Given two dead states \(s ∈ S\) and \(u ∈ U\), then the singleton bisimulation \({∼} ⊆ \norm(S) × \norm(U)\):
    \begin{mathpar}
        {∼} ≜ \{(s,u)\} \and 
        δ_{∼}((s,u), α) ≜ \reject
    \end{mathpar}
    is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
\end{lemma}

\begin{proof}
    By computation
\end{proof}

\begin{lemma}\label{thm:recursive-construction-lemma}
    Given two GKAT coalgebra \(S\) and \(U\), and two of their elements \(s ∈ S\) and \(u ∈ U\),
    there exists a bisimulation \({∼} ⊆ \norm(⟨s⟩) × \norm(⟨u⟩)\) s.t. \(s ∼ u\), if and only if there exists a normalized bisimulation \({≃} ⊆ ⟨s⟩ × ⟨u⟩\) s.t. \(s ≃ u\).
\end{lemma}

\begin{proof}
    Recall that there exists a bisimulation \({∼} ⊆ \norm(⟨s⟩) × \norm(⟨u⟩)\) if and only if for all \(s₁ ∼ u₁\):
    \begin{itemize}
        \item for all results \(r ∈ \{\accept, \reject\}\): \(δ_{\norm(S)}(s₁, α) = r ⟺ δ_{\norm(U)}(u₂, α) = r\);
        \item otherwise, let \((s₂, p) ≜ δ_{\norm(S)}(s₁, α)\) and \((u₂, q) ≜ δ_{\norm(u)}(u₁, α)\), then \(p = q\) and \(s₂ ∼ u₂\)
    \end{itemize}

    We prove \(⟹\) direction by constructing the following normalized bisimulation from the bisimulation \({∼}\):
    \[{≃} ≜ {∼} ∪ \{(s, u) ∣ \text{both \(s ∈ S, u ∈ U\) are dead}\}.\]
    We then show that \({≃}\) satisfies all the condition in~\Cref{def:norm-bism}.

    The condition \labelcref{itm:acc-condition} holds if \(s ∼ u\):
    \begin{align*}
        δ_S(s, α) = \accept 
        ⟺{}& δ_{\norm(S)}(s, α) = \accept \\
        ⟺{}& δ_∼((s,u), α) = \accept \\ 
        ⟺{}& δ_{\norm(U)}(u, α) = \accept \\
        ⟺{}& δ_U(u, α) = \accept
    \end{align*}
    The condition \labelcref{itm:acc-condition} also holds if \(s, u\) are both dead, because both can never accepts any atoms.

@@

    The condition~\labelcref{itm:transition-bisim} holds when \(s ∼ u\).
    First note that \(s'\) and \(u'\) has to be both live or both dead: because \(δ_S(s, α) = (s', p)\), then \(\norm(δ_S)(s', α)\) can either be rejection or \((s',p)\), and so is \(\norm(δ_U)(u', α)\):
    \begin{align*}
        s' \text{ is live} 
        & ⟺ δ_{\norm(S)}(s, α) = (s', p) \\
        & ⟺ δ_{\norm(U)}(u, α) = (u', p) \\
        & ⟺ u' \text{ is live}.
    \end{align*}
    \begin{itemize}
        \item If both \(s'\) and \(u'\) are live, then \(s' ∼ u'\). By~\cref{thm:sub-coalg-preserve-bisim}, the bisimulation \(∼'\) is just \(∼\) restricted to \(⟨s'⟩\) and \(⟨u'⟩\).
        \item If both \(s'\) and \(u'\) are dead, then \(∼'\) can just be the singleton relation, according to~\cref{thm:bisim-between-dead}.
    \end{itemize}

    The condition~\labelcref{itm:transition-dead} holds: by the proof of condition~\labelcref{itm:transition-bisim}, \(s'\) and \(u'\) has to be either both live or both dead; if they are both live, then there cannot be a element in \(G(∼)\) that can project to \((s', p)\) under \(π₁\) but projects to \((t', q)\) under \(π₂\). Thus both \(s'\) and \(t'\) has to be dead.

    We then show the \(⟸\) direction, for arbitrary \(s' ∈ S\) and \(u' ∈ U\), we use \(≡_{s', u'}\) to denote the maximal bisimulation between \(\norm(⟨s'⟩)\) and \(\norm(⟨u'⟩)\).
    \begin{align*}
        {∼'} ≜ ⋃ \{≡_{s', u'} & ∣ ∃ α ∈ \At, p ∈ K, \\*
            & δ_{\norm(S)}(s, α) = (s', p) \\*
            & \text{ and } δ_{\norm(U)}(u, α) = (u', p)\}.
    \end{align*}
    For all the \(s'\) and \(u'\) in the above definition, \(⟨s'⟩ ⊑ ⟨s⟩\) and \(⟨u'⟩ ⊑ ⟨u⟩\), therefore by~\Cref{thm:norm-sub-coalg}, \(\norm(⟨s'⟩) ⊑ \norm(⟨s⟩)\) and \(\norm(⟨u'⟩) ⊑ \norm(⟨u⟩)\). 
    By~\Cref{thm:sub-coalg-preserve-bisim}, every \(≡_{s', u'}\) is a bisimulation between \(\norm(⟨s⟩)\) and \(\norm(⟨t⟩)\), and because bisimulation is closed under arbitrary union~\cite{rutten_UniversalCoalgebraTheory_2000}, \(∼'\) is a bisimulation between \(\norm(⟨s⟩)\) and \(\norm(⟨t⟩)\).

    To obtain the desired bisimulation \({∼}\) between \(\norm(⟨s⟩)\) and \(\norm(⟨u⟩)\), we add the pair \((s, t)\) to \(∼'\), 
    \[{∼} ≜ {∼'} ∪ \{(s, u)\},\] 
    with the following transition \(δ_{∼}\): for all \(α ∈ \At\),
    \begin{itemize}[nosep]
        \item if \(δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \accept\), then \(δ_{∼}((s, u), α) ≜ \accept\);
        \item if \(δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \reject\), then \(δ_{∼}((s, u), α) ≜ \reject\);
        \item if \(δ_{\norm(S)}(s, α) = (s', p)\) and \(δ_{\norm(U)}(u, α) = (u', p)\), then \(δ_{∼}((s, u), α) = ((s', u'), p)\);
        \item for all \((s', u') ∈ {∼}'\) that is not equal to \((s, u)\), we let \(δ_∼\) inherits the transition of \(δ_{∼'}\), i.e. \(δ_{∼}((s', u'), α) = δ_{∼'}((s', u'), α)\)
    \end{itemize}
    \emph{if \(δ_{∼}\) is well-defined}, then we can verify that \(∼\) is indeed a bisimulation between \(\norm(⟨s⟩)\) and \(\norm(⟨u⟩)\) where \(s ∼ u\). 
    We show the slightly more complicated case: \(δ_{\norm(S)}(s, α) = (s', p)\) and \(δ_{\norm(U)}(u, α) = (u', p)\) implies \(s' ∼ u'\) as an example. 
    By condition~\labelcref{itm:transition-bisim}, there exists a bisimulation \(∼_{s', u'} ⊆ \norm(⟨s'⟩) × \norm(⟨u'⟩)\), and because \(≡_{s', u'}\) is the maximal bisimulation between \(\norm(⟨s'⟩)\) and \(\norm(⟨u'⟩)\),
    \[(s', u') ∈ {∼_{s', u'}} ⊆ {≡_{s', u'}} ⊆ {∼}.\]

    Finally, we demonstrate that \(δ_∼\) is well-defined by leveraging the conditions in~\Cref{thm:recursive-construction}. 
    Specifically, we will show that the definition of \(δ_∼\) covers all the possible cases, by case analysis on the result of \(δ_S\): for all \(α ∈ \At\),
    \begin{itemize}
        \item If \(δ_S(s, α) = \accept\), then by condition~\labelcref{itm:acc-condition}, \(δ_U(u, α) = \accept\); therefore \[δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \accept.\]
        \item If \(δ_S(s, α)\) transitions to a dead state or reject, then by condition~\labelcref{itm:rej-or-dead} \(δ_U(u, α)\) will also transition to a dead state or reject, then \[δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \reject.\]
        \item If \(δ_S(s, α) = (s', p)\) and \(s'\) is live, then \(δ_U(u, α) = (u', p)\) necessarily holds, otherwise it would violate one of conditions~\labelcref{itm:acc-condition,itm:rej-or-dead,itm:transition-dead}.

        By condition~\labelcref{itm:transition-bisim}, there exists a bisimulation \(∼_{s', u'}\) between \(\norm(⟨s'⟩)\) and \(\norm(⟨u'⟩)\) s.t. \(s' ∼_{s', u'} u'\). Because bisimulation preserves liveness (\Cref{thm:bisim-preserve-liveness}), \(s', u'\) has to be both dead or live. 
        Finally, because \(s'\) is live, therefore \(u'\) is also live, and we obtain the final case in the definition of \(δ_{∼}\): \(δ_{\norm(S)}(s, α) = (s', p)\) and \(δ_{\norm(U)}(u, α) = (u', p)\).
        \qedhere
    \end{itemize}
\end{proof}

With some slight tweak to \Cref{thm:recursive-construction-lemma}, it will be applicable to trace equivalence instead of bisimulation.
The correctness of this modification is supported by the correspondence between trace equivalence and the existence of a bisimulation between normalized GKAT coalgebras, which is stated in~\Cref{thm:norm-bisim-correctness}.

\begin{theorem}[Recursive Construction]\label{thm:recursive-construction}
    For any two states in two GKAT coalgebra \(s ∈ S, u ∈ U\), \(s\) and \(u\) are finite-trace equivalent \(⟦s⟧_{S} = ⟦u⟧_{U}\) if and only if all the following conditions hold:
    \begin{enumerate}
        \item for all \(α ∈ \At\), \(δ_{S}(s, α) = \accept ⟺ δ_{U}(u, α) = \accept\);
        \item \(s\) reject \(α\) or transition to a dead state via \(α\) if and only if \(u\) rejects \(α\) or transition to a dead state via \(α\);  
        \item if \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', p)\), then \(⟦s'⟧_{S} = ⟦u'⟧_{U}\);
        \item if \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(t'\) are dead.
    \end{enumerate}
    Notice that above condition is similar to those in~\cref{thm:recursive-construction}, except bisimulation of \(s'\) and \(u'\) is replaced with trace equivalence.
\end{theorem}

\begin{proof}
    By the standard argument with normalization preserves sub-coalgebra (\Cref{thm:norm-sub-coalg}), we can obtain for all \(s ∈ S\) and \(u ∈ U\), \(\norm(⟨s⟩) ⊑ \norm(S)\) and \(\norm(⟨u⟩) ⊑ \norm(U)\).
    Therefore, because subcoalgebra preserve and reflect bisimulation (\Cref{thm:sub-coalg-preserve-bisim}) and the correctness of bisimulation on normalized coalgebra (\Cref{thm:norm-bisim-correctness}): for all \(s ∈ S\) and \(u ∈ U\),
    \begin{align*}
        & ∃~ {∼} ⊆ \norm(⟨s⟩) × \norm(⟨u⟩), s ∼ u \\
        ⟺{}& ∃~ {∼} ⊆ \norm(S) × \norm(U), s ∼ u \\  
        ⟺{}& ⟦s⟧_S = ⟦u⟧_U.
    \end{align*}
    We can instantiate the above equivalence to \(s', u'\) and \(s, u\) respectively: the instantiation to \(s', u'\) will give us the conditions in this theorem is equivalent to the ones in~\Cref{thm:recursive-construction-lemma}; and instantiation to \(s, u\) show that existence of bisimulation is equivalent to trace equivalence:
    \begin{align*}
        & \text{Conditions in \Cref{thm:recursive-construction-lemma} hold} \\  
        ⟺{}& \text{Conditions in the current theorem hold}\\
        ⟺{}& ∃~ {∼} ⊆ \norm(⟨s⟩) × \norm(⟨u⟩), s ∼ u\\
        ⟺{}& ⟦s⟧_{S} = ⟦u⟧_{U}. \qedhere
    \end{align*}
\end{proof}

The above construction theorem already gives us an algorithm to recursively decide whether \(⟦s⟧_{S} = ⟦u⟧_U\), when \(⟨s⟩_S\) and \(⟨u⟩_U\) is finite.
However, this algorithm can be further optimized: we will derive that a dead state is only trace equivalent to other dead states. 
This means that when checking the trace-equivalence of states \(s\) and \(u\), if we already know one of them is dead, we only need to check whether the other is dead, instead of going through all the conditions in~\Cref{thm:recursive-construction}.

\begin{theorem}\label{thm:bisim-one-dead}
    Given two states \(s ∈ S\) and \(u ∈ U\), if \(s\) is a dead state in \(S\), then \(s\) and \(u\) is trace equivalent if and only if \(u\) is also dead.
\end{theorem}

\begin{proof}
    TODO: I think we can refine \(⟹\) direction, but I am not sure...

>>>>

\newcommand\altxrightarrow[2][0pt]{\mathrel{\ensurestackMath{\stackengine%
  {\dimexpr#1-7.5pt}{\xrightarrow{\phantom{#2}}}{\scriptstyle\!#2\,}%
  {O}{c}{F}{F}{S}}}}
\newcommand{\transvia}[1]{
    \mathrel{\raisebox{-2px}{\(\altxrightarrow[-2px]{#1}\)}}
}
\newcommand{\transAcc}[2]{⇒_{#1} #2}
\newcommand{\transRej}[2]{↓_{#1} #2}

@@

\begin{theorem}[normalization perserves liveness]\label{thm:norm-perserve-liveness}
    A state \(s ∈ S\) is dead (live) in \(S\) if and only if it is dead (live) in \(\norm(S)\).
\end{theorem}

@@

\begin{definition}[normalizing bisimulation]\label{def:norm-bisim}
    Given two GKAT coalgebra \(S\) and \(U\), and a bisimulation \({∼} ⊆ \norm(S) × \norm(U)\), a normalizing bisimulation \({≈} ⊆ S × U\) is a GKAT coalgebra that contains \(∼\) with the following transition function:
    \[
        δ_≈((s, u), α) ≜ \begin{cases}
            \accept & \begin{aligned}
                & \text{if } δ_S(s, α) = \accept \\[-3px]
                & \text{and } δ_U(s, α) = \accept
            \end{aligned}\\[5px]
            ((s', u'), p) & 
                \begin{aligned}
                    & \text{if } s \transvia{α ∣ p}_{S} s' \\[-3px]
                    & \text{and } u \transvia{α ∣ p}_{U} u' 
                \end{aligned}\\[5px]
            \reject & \text{otherwise}
        \end{cases}
    \]
    We call \(∼\) \emph{the underlying bisimulation} of \(≈\).
\end{definition}

\begin{lemma}
    Given a bisimulation \({∼} ⊆ \norm(S) × \norm(U)\), the following relation \({∼_d} ⊆ S × U\):
    \[{∼_d} ≜ {∼} ∪ \{(s, u) ∣ \text{\(s\) and \(u\) are dead in \(S\) and \(U\)}\},\]
    form a GKAT coalgebra the transition function \(δ_≈\) in~\Cref{def:norm-bisim}
\end{lemma}

\begin{proof}
    We will need to show \(∼_d\) is closed under the transition of \(δ_{≈}\) in~\Cref{def:norm-bisim}.
    Specifically, if \(s ∼_d u\) and \(s \transvia{α ∣ p}_{S} s'\) and \(u \transvia{α ∣ p} u'\), then \(s' ∼_d u'\).
    By case analysis on the liveness of \(s'\) and \(u'\) in \(S\) and \(U\):
    \begin{itemize}
        \item If both \(s'\) and \(u'\) are dead, then \(s ∼_d u\) by definition of \(∼_d\).
        \item If both \(s'\) and \(u'\) are live, then \(s, s'\) and \(u, u'\) are live in \(S\) and \(U\) respectively. 
        Thus, \(s \transvia{α ∣ p}_{\norm(S)} s'\), and \(u \transvia{α ∣ p}_{\norm(U)} u'\).
        Therefore, \(s' ∼ u'\) by definition of bisimulation on \(\norm(S)\) and \(\norm(U)\); and \(s ∼_d u'\).
        \item If \(s'\) is live but \(u'\) is dead, then \(s\) is live, and \(s ∼ u\).
        However, \(s \transvia{α ∣ p}_{\norm(S)} s'\) yet \(u \transRej{\norm(U)}{u}\), which violates the condition for bisimulation, thus this case cannot happen.
        \item If \(u'\) is live but \(s'\) is dead, this case also cannot happen similar to the previous case.
    \end{itemize}
\end{proof}

\begin{lemma}\label{thm:norm-bisim-exists-unique}
    For any \(∼\) a normalizing bisimulation with \(∼\) as underlying bisimulation exists and is unique.
\end{lemma}

\begin{proof}
    TODO: sub-coalgebra form a complete lattice.
\end{proof}

\begin{lemma}\label{thm:norm-bisim-not-bisim-then-component-dead}
    Every state in \((s, u) ∈ {≈} ⊆ S × U\) but not in \({∼} ⊆ \norm(S) × \norm(U)\) then \(s\) is dead in \(S\), \(u\) is dead in \(U\).
\end{lemma}

\begin{proof}
    Because \(∼_d\) is a GKAT coalgebra with transition \(δ_≈\) and also contains all the states in \(∼\), therefore \({≈} ⊑ {∼_d}\).
    Thus, for all \((s, u) ∈ {≈}\) but not in \(∼\), \(s\) and \(u\) must be dead in \(S\) and \(U\).
\end{proof}

\begin{lemma}\label{thm:live-in-component-then-live-in-norm-bisim}
    Given a normalizing bisimulation \(≈\), if a state \((s, u)\) is live in \(≈\), then \(s\) and \(u\) are live in \(S\) and \(U\) respectively, which also implies that \((s, u)\) is live in \(∼\).
\end{lemma}

\begin{proof}
    TODO: by preservation of accepting state and liveness chain.
\end{proof}

\begin{lemma}
    Every state in \((s, u) ∈ {≈} ⊆ S × U\) but not in \({∼} ⊆ \norm(S) × \norm(U)\) then \((s, u)\) is dead in \(≈\).
\end{lemma}

\begin{proof}
    By~\Cref{thm:norm-bisim-not-bisim-then-component-dead}, \(s\) is dead in \(S\), then by contrapositive of~\Cref{thm:live-in-component-then-live-in-norm-bisim}, we obtain that \((s, u)\) is dead in \(≈\).
\end{proof}

\begin{theorem}[Correctness]
    Given two states in GKAT coalgebras \(s ∈ S\) and \(u ∈ ,U\), there exists a normalizing bisimulation \({≈} ⊆ S × U\) s.t. \(s ≈ u\), if and only if \(⟦s⟧_S = ⟦u⟧_U\).
\end{theorem}

\begin{proof}
    \(⟹\) direction, because normalization perserves carrier set, therefore \((s, u) ∈ \norm(≈)\).
    And because \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\), \(⟦s⟧_S = ⟦u⟧_U\).

    \(⟸\) direction, because \(⟦s⟧_S = ⟦u⟧_U\), then by~\cref{thm:norm-bisim-correctness}, then there a bisimulation \(∼ ⊆ \norm(S) × \norm(U)\) s.t. \(s ∼ u\). 
    Because~\Cref{thm:norm-bisim-exists-unique}, therefore there exists a normalizing bisimulation \(≈\) with \(∼\) as its underlying bisimulation. 
    Because the carriers \({∼} ⊆ {≈}\), therefore \(s ∼ u\) implies \(s ≈ u\).
\end{proof}

\begin{theorem}[Recursive Construction]
    Given a relation \({≈} ⊆ S × U\), \(≈\) is a carrier of some normalizing bisimulation if and only if all the following condition are satisfied: for all pair of states \(s ≈ u\),
    \begin{enumerate}
        \item\label{itm:acc-condition} for all \(α ∈ \At\), 
        \(s \transAcc{S}{α}\) if and only if \(u \transAcc{U}{α}\);
        \item\label{itm:rej-or-dead} if \(s\) reject \(α\) but \(u\) transitions to \(u'\), then \(u'\) is dead; similarly when \(u\) rejects \(α\);  
        \item\label{itm:transition-bisim} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', p)\), then \(s' ≈ u'\).
        \item\label{itm:transition-dead} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(u'\) are dead.
    \end{enumerate}
\end{theorem}

\begin{proof}
    To show the \(⟹\) direction, assuming there exists a normalizing bisimulation \(≈\), then it would satisfy all the conditions. 
    We first note that the pairs \((s, u) ∈ {≈}\) can satisfy one of two properties (or both), either \(s\) and \(u\) are both dead in \(S\) and \(U\), or \((s, u)\) is in the underlying bisimulation \(∼\).
    Then we will verify that all the condition is satisfied both when \(s ∼ u\) and when \(s, u\) are dead.

    The condition~\labelcref{itm:acc-condition} holds if \(s ∼ u\):
    \begin{align*}
        s \transAcc{S}{α}
        \text{ iff }& s \transAcc{\norm(S)}{α} \\
        \text{ iff }& (s, u) \transAcc{∼}{α} \\ 
        \text{ iff }& u \transAcc{\norm(U)}{α} \\
        \text{ iff }& u \transAcc{U}{α};
    \end{align*}
    condition~\labelcref{itm:acc-condition} also holds if \(s\) and \(u\) are dead in \(S\) and \(U\), because both can never accept any atoms.

@@

    The condition~\labelcref{itm:transition-bisim} holds because when \(s \transvia{α ∣ p}_S s'\) and \(u \transvia{α ∣ p}_U u'\), then \((s, u) \transvia{α ∣ p}_{≈} (s', u')\) by definition of normalizing bisimulation (\Cref{def:norm-bisim}).

    The condition~\labelcref{itm:transition-dead} holds, let \(δ_S(s, α) = (s', p)\) and \(δ_U(u, α) = (u', q)\) and \(p ≠ q\). 
    \begin{itemize}
        \item If \(s ∼ u\), then we can show condition~\labelcref{itm:transition-dead} by case analysis on the liveness of \(s'\) and \(u'\):
        if \(s', u'\) are not both dead, then the condition of \({∼} ⊆ \norm(S) × \norm(U)\) cannot be satisfied, hence \((s', u') ∉ {∼}\), which by~\Cref{thm:norm-bisim-not-bisim-then-component-dead} means that \(s'\) and \(u'\) are both dead, contradicting the premise.
        Therefore, \(s'\) and \(u'\) needs to be both dead.
        \item If \(s\) and \(u\) are both dead, then by~\Cref{thm:dead-iff-all-reachable-dead}, both \(s\) and \(u\) can only transition to dead state, hence \(s'\) and \(u'\) are both dead.
    \end{itemize}

    Then we prove the \(⟸\) direction.
    TODO: by constructing a \(≈\) and show that \(\norm(≈)\) is a bisimulation. Maybe using the maximal bisimulation?
\end{proof}

\begin{lemma}\label{thm:bisim-between-dead}
    Given any two dead states \(s ∈ S\) and \(u ∈ U\), and the maximal bisimulation \({≡} ⊆ \norm(S) × \norm(U)\), then we have \(s ≡ u\).
\end{lemma}

\begin{theorem}[Equivalence Relation]
    Given two states in GKAT coalgebra \(s ∈ S\) and \(u ∈ U\), there exists a normalizing bisimulation \(≈\) between \(S\) and \(U\) s.t. \(s ≈ u\), if and only if there exists a normalizing bisimulation that is an \emph{equivalence relation} \(≊\) in \(S + U\) s.t. \(s ≊ u\).
\end{theorem}

\begin{proof}
    First, recall that by~\Cref{thm:} \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
    Because \(S ⊑ S + U\) and \(U ⊑ S + U\) and normalization preserves sub-coalgebra~\Cref{thm:norm-sub-coalg}, we have \(\norm(S) ⊑ \norm(S + U)\) and \(\norm(U) ⊑ \norm(S + U)\); and by~\Cref{thm:sub-coalg-preserve-liveness}, \(s ∈ S\) is dead in \(S + U\) if and only if it is dead in \(S\), and similarly for \(u ∈ U\).

    The \(⟹\) direction. 
    Because sub-coalgebra preserves bisimulation (\Cref{thm:sub-coalg-preserve-bisim}), then \(\norm(≈)\) is a bisimulation between \(\norm(S + U)\) and itself.
    We consider \(≡\) to be the maximal bisimulation between \(\norm(S + U)\) and itself, then \(≡\) is an equivalence relation~\cite[]{}.

    Let \(≊\) be the normalizing bisimulation generated by \(≡\), then \(≊\) has the same carrier as \(≡\).
    All the \((s, u) ∈ {≊}\) are either \(s ≡ u\) or \(s\) and \(u\) are dead in \(S\) and \(U\), which implies that both are dead in \(S + U\).
    However, \(≡\) already contains all the dead pairs (\cref{thm:bisim-between-dead}), therefore for all \((s, u) ∈ {≊}\), they are also in \(≡\), implying \({≡} ⊇ {≊}\).
    Finally, by definition of normalizing bisimulation (\Cref{def:norm-bisim}), \({≡} ⊆ {≊}\).
    Therefore, \(≡\) and \(≊\) share the same carrier and \(≊\) is a normalizing bisimulation that is also an equivalence relation, which also contains \((s, u)\):
    \[(s, u) ∈ {≈} = \norm(≈) ⊆ {≡} ⊆ {≊},\]
    where all the \(=\) and \(⊆\) are orders on the carrier set, instead of the coalgebra.

    The \(⟸\) direction can be split into two cases, by whether \((s, u)\) is in the underlying bisimulation of \(≊\), which we denote as \(≃\): 
    \begin{itemize}
        \item If \(s ≃ u\), then we can reflect \(≃\)  through \(\norm(S) ⊑ \norm(S + U)\) and \(\norm(U) ⊑ \norm(S + U)\) (\Cref{thm:sub-coalg-preserve-bisim}) to obtain a bisimulation \(∼ ⊆ \norm(S) × \norm(U)\) and \(s ∼ u\).
        Taking the normalizing bisimulation over \(∼\) will yield a desired normalizing bisimulation \(≈\) s.t. \(s ≈ u\).
        \item If \((s, u)\) are not in \(≃\), then they are both dead in \(S + U\) (\Cref{thm:norm-bisim-not-bisim-then-component-dead}), which implies \(s\) and \(u\) are respectively dead in \(S\) and \(U\) (\Cref{thm:sub-coalg-preserve-liveness}).
        Then we take \(≈\) to be normalizing bisimulation generated form the maximal bisimulation between \(S\) and \(U\), which contains \(s\) and \(u\) (\Cref{thm:bisim-between-dead}).
        \qedhere
    \end{itemize}
\end{proof}

% % \begin{definition}[Normalizable Bisimulation]
% %     A relation \({≃} ⊆ S × U\) is a normalizable bisimulation if it has a GKAT coalgebra structure \(δ_≃\), and \(\norm(≃)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
% % \end{definition}

% % \begin{lemma}[Preservation of Liveness]
% %     For any normalizable bisimulation \({≃} ⊆ S × U\) if \(s ≃ u\), then \(s\) is live in \(S\) if and only if \(u\) is live in \(U\).
% % \end{lemma}

% % \begin{proof}
% %     Because \(\norm(≃)\) has two projection homomorphism to \(\norm(S)\) and \(\norm(U)\), then 
% % \end{proof}

% % \begin{theorem}[Correctness]
% %     If there exists a normalizable bisimulation \({≃} ⊆ S × U\) s.t. \(s ∼ u\) for \(s ∈ S\) and \(u ∈ U\), then \(⟦s⟧_S = ⟦u⟧_U\).
% % \end{theorem}

% % \begin{theorem}[Construction]
% %     Given two states \(s ∈ S, u ∈ U\) then there exists a normalizable bisimulation \({≃'} ⊆ S × U\) if and only if there exists a relation \({≃} ⊆ S × U\) that satisfy the following conditions:
% %     \begin{enumerate}
% %         \item\label{itm:acc-condition} for all \(α ∈ \At\), \(δ_{S}(s, α) = \accept ⟺ δ_{U}(u, α) = \accept\);
% %         \item\label{itm:rej-or-dead} if \(s\) reject \(α\) but \(u\) transitions to \(u'\), then \(u'\) is dead; similarly when \(u\) rejects \(α\);  
% %         \item\label{itm:transition-bisim} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', p)\), then \(s' ≃ u'\).
% %         \item\label{itm:transition-dead} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(t'\) are dead.
% %     \end{enumerate}
% % \end{theorem}

% % \begin{proof}
% %     The \(⟸\) direction can be shown that 
% % \end{proof}

% % \begin{definition}[normalizing bisimulation]\label{def:norm-bism}
% %     A relation \({≃} ⊆ S × U\) on two GKAT coalgebra \(S\) and \(U\) is a normalizing bisimulation if and only if for any two states \(s ≃ u\), all the following holds:
% %     \begin{enumerate}
% %         \item\label{itm:acc-condition} for all \(α ∈ \At\), \(δ_{S}(s, α) = \accept ⟺ δ_{U}(u, α) = \accept\);
% %         \item\label{itm:rej-or-dead} if \(s\) reject \(α\) but \(u\) transitions to \(u'\), then \(u'\) is dead; similarly when \(u\) rejects \(α\);  
% %         \item\label{itm:transition-bisim} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', p)\), then \(s' ≃ u'\).
% %         \item\label{itm:transition-dead} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(t'\) are dead.
% %     \end{enumerate}
% %     And \(≃\) can be endowed with the following GKAT coalgebra structure:
% %     \[
% %         δ_≃((s, u), α) ≜ \begin{cases}
% %             \accept & \begin{aligned}
% %                 & \text{if } δ_S(s, α) = \accept \\[-3px]
% %                 & \text{and } δ_U(s, α) = \accept
% %             \end{aligned}\\[5px]
% %             ((s', u'), p) & 
% %                 \begin{aligned}
% %                     & \text{if } s \transvia{α ∣ p}_{S} s' \\[-3px]
% %                     & \text{and } u \transvia{α ∣ p}_{U} u' 
% %                 \end{aligned}\\[5px]
% %             \reject & \text{otherwise}
% %         \end{cases}
% %     \]
% % \end{definition}

% \begin{lemma}[Perservation of Liveness]\label{thm:norm-bisim-perserve-liveness}
%     For any normalizing bisimulation \(≃\) and a pair of states \(s ≃ u\), \((s, u)\) is live in \(≃\) if and only if \(s\) is live in \(S\), and similarly for \(u\) in \(U\). 
%     This implies that if \(s ≃ u\), then \(s\) is live in \(S\) if and only if \(u\) is live in \(U\).
% \end{lemma}

% \begin{proof}
%     First by definition of normalizing bisimulation that a state \(s\) is accepting in \(S\) if and only if \((s, u)\) is accepting in \(≃\).

%     Then, to prove the preservation of liveness, we prove the preservation of live chain, i.e. there exists a chain to accepting state in \(≃\): \((s₀, u₀) \transvia{α₁ ∣ p₁} (s₁, u₁) \transvia{α₂ ∣ p₂} (s₂, u₂) \transvia{α₂ ∣ p₂} ⋯ \transvia{αₙ ∣ pₙ} (sₙ, uₙ)\) if and only if there exists a similar accepting chain in \(S\): \(s₀ \transvia{α₁ ∣ p₁} s₁ \transvia{α₂ ∣ p₂} s₂ \transvia{α₂ ∣ p₂} ⋯ \transvia{αₙ ∣ pₙ} sₙ\).

%     Because every state on an accepting chain is live, all we need to prove is that for two states \(s ≃ u\),
%     \begin{enumerate}
%         \item if \((s', u')\) is a live state in \(≃\) and \((s, u) \transvia{α ∣ p}_≃ (s', u')\), then \(s \transvia{α ∣ p}_S s'\)
%         \item if \(s'\) is a live state in \(S\) and \(s \transvia{α ∣ p} s'\), then \((s, u) \transvia{α ∣ p} (s', u')\) for some \(u'\).
%     \end{enumerate}
%     The first goal can be shown by definition of \(δ_{≃}\). 
%     Because \(s'\) in the second goal is live, therefore only condition~\labelcref{itm:transition-bisim} in~\Cref{def:norm-bism} matches the condition, thus implying \(u \transvia{α ∣ p} u'\) and \((s, u) \transvia{α ∣ p} (s', u')\).
% \end{proof}

% \begin{theorem}
%     Given a normalizing bisimulation \(≃: S × U\), its normalization is indeed a bisimulation between \(\norm(S)\) and \(\norm(U)\)
% \end{theorem}

% \begin{proof}
%     We will need to prove the two projection of \(π₁: {≃} → S\) and \(π₂: {≃} → U\) is a homomorphism, without loss of generality, we will only show that \(π₁\) is a homomorphism.

%     First the accepting case, for all \(s ≃ u\):
%     \begin{align*}
%         δ_{\norm(≃)}((s, u), α) = \accept
%         & ⟹ δ_{≃}((s, u), α) = \accept \\
%         & ⟹ δ_{S}(s, α) = \accept \\ 
%         & ⟹ δ_{\norm(S)}(s, α) = \accept \\ 
%     \end{align*}
